Sparse spectral and $p$-finite element methods on disk-slices and trapeziums I: zero boundary conditions

### 2019-04-03 papers/halfdisk/halfdisk.tex
Further edits

Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Define, for $n=0,1,2,\dots$: 
\begin{align}
\bigPab_n := \begin{pmatrix}
		P^{(a,b)}_{n,-n}(x,y) \\
		\vdots \\
		P^{(a,b)}_{n,n}(x,y)
	\end{pmatrix} \in \R^{n+1}, 
\quad \quad 
\bigPab := \begin{pmatrix}
		\bigPab_0 \\
		\bigPab_1 \\
		\bigPab_2 \\
		\vdots \\
	\end{pmatrix}
\end{align}
and set $J_x^{(a,b)}, J_y^{(a,b)}$ as the Jacobi matrices corresponding to
\begin{align}
J_x^{(a,b)} \: \bigPab(x,y) = x \: \bigPab(x,y), \quad J_y^{(a,b)} \: \bigPab(x,y) = y \: \bigPab(x,y).
\label{eqn:jacobimatricesdefinition}
\end{align}
Then $J_x^{(a,b)}, J_y^{(a,b)}$ have the Banded-Block-Banded structures
\begin{align}
J_{x/y}^{(a,b)} &= \begin{pmatrix}
		B^{x/y}_0 & A^{x/y}_0 & & & & \\
		C^{x/y}_1 & B^{x/y}_1 & A^{x/y}_1 & & & \\
		& C^{x/y}_2 & B^{x/y}_2 & A^{x/y}_2  & & & \\
		& & C^{x/y}_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{pmatrix}
\end{align}
where
\begin{align}
A^x_n &:= \begin{pmatrix}
		\alphaab_{n+1,0,1} & 0 & \hdots & 0 \\
		& \ddots & & \vdots & \\
		& & \alphaab_{n+1,n,1} & 0 \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
B^x_n &:= \begin{pmatrix}
		\alphaab_{n,0,2} & & \\
		& \ddots & \\
		& & \alphaab_{n,n,2} \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots \\
C^x_n &:= \big( A^x_n \big)^T \in \R^{(n+1)\times n},  \quad n = 1,2,\dots \\ 
\nonumber \\
A^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,6} & & & \\
		\betaab_{n,1,5} & 0 & & & \\
		& \ddots & \ddots & \ddots & \\
		& & \betaab_{n,n,5}& 0 & \betaab_{n,n,6} \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
B^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,4} & & \\
		\betaab_{n,1,3} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-1,4} \\
		& & \betaab_{n,n,3} & 0
	    \end{pmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots \\
C^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,2} & & \\
		\betaab_{n,1,1} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-2,2} \\
		& & \ddots & 0 \\
		& & & \betaab_{n,n,1} \\
	    \end{pmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots
\end{align}

@@

Let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bigPab(x,y)^\top \mathbf{f}$. Recall that $\Wab(x,y) := w_H^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)})$, where $\rho(x) := (1-x^2)^\half$, $w_H^{(a,b)}(x) := x^a \: (1-x^2)^b$ and $w_P^{(b)}(y) := (1-y^2)^b$ for the half disk case.

@@

We can gain the non-zero entries of the weighted differential operators similarly, by noting that
\begin{align}
	\pddx \Wab(x,y) &= a W^{(a-1, b)}(x,y) - 2bW^{(a+1, b-1)}(x,y) \\
	\pddy \Wab(x,y) &= -2b \: y \: W^{(a, b-1)}(x,y)
\end{align}
and also that
\begin{align}
	\ip< \Wab \Pnkab, \Wab \Pab_{m,j} >_{W^{(-a, -b)}} = \ip< \Pnkab, \Pab_{m,j} >_\Wab = \delta_{n,m} \: \delta_{k,j}.
\end{align}

>>>>

Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Define, for $n=0,1,2,\dots$: 
\begin{align}
\bigPab_n := \begin{pmatrix}
		P^{(a,b)}_{n,-n}(x,y) \\
		\vdots \\
		P^{(a,b)}_{n,n}(x,y)
	\end{pmatrix} \in \R^{n+1}, 
\quad \quad 
\bigPab := \begin{pmatrix}
		\bigPab_0 \\
		\bigPab_1 \\
		\bigPab_2 \\
		\vdots \\
	\end{pmatrix}
\end{align}
and set $J_x^{(a,b)}, J_y^{(a,b)}$ as the Jacobi matrices corresponding to
\begin{align}
J_x^{(a,b)} \: \bigPab(x,y) = x \: \bigPab(x,y), \quad J_y^{(a,b)} \: \bigPab(x,y) = y \: \bigPab(x,y).
\label{eqn:jacobimatricesdefinition}
\end{align}
The matrices $J_x^{(a,b)}, J_y^{(a,b)}$ act on the coefficients vector of a functions expansion in the $\{\Pnkab\}$ basis. For example, let $a, b$ be general parameters and a function $f(x,y)$ defined on $\Omega$ be approximated by its expansion $f(x,y) = \bigPab(x,y)^\top \mathbf{f}$. Then $x \: f(x,y)$ is approximated by $\bigPab(x,y)^\top {J_x^{(a,b)}}^\top \mathbf{f}$. In other words, ${J_x^{(a,b)}}^\top \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y) \to x \: f(x,y)$ in the  $\{\Pnkab\}$ basis.

$J_x^{(a,b)}, J_y^{(a,b)}$ have the Banded-Block-Banded structures
\begin{align}
J_{x/y}^{(a,b)} &= \begin{pmatrix}
		B^{x/y}_0 & A^{x/y}_0 & & & & \\
		C^{x/y}_1 & B^{x/y}_1 & A^{x/y}_1 & & & \\
		& C^{x/y}_2 & B^{x/y}_2 & A^{x/y}_2  & & & \\
		& & C^{x/y}_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{pmatrix}
\end{align}
where
\begin{align}
A^x_n &:= \begin{pmatrix}
		\alphaab_{n+1,0,1} & 0 & \hdots & 0 \\
		& \ddots & & \vdots & \\
		& & \alphaab_{n+1,n,1} & 0 \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
B^x_n &:= \begin{pmatrix}
		\alphaab_{n,0,2} & & \\
		& \ddots & \\
		& & \alphaab_{n,n,2} \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots \\
C^x_n &:= \big( A^x_n \big)^T \in \R^{(n+1)\times n},  \quad n = 1,2,\dots \\ 
\nonumber \\
A^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,6} & & & \\
		\betaab_{n,1,5} & 0 & & & \\
		& \ddots & \ddots & \ddots & \\
		& & \betaab_{n,n,5}& 0 & \betaab_{n,n,6} \\
	    \end{pmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots \\
B^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,4} & & \\
		\betaab_{n,1,3} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-1,4} \\
		& & \betaab_{n,n,3} & 0
	    \end{pmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots \\
C^y_n &:= \begin{pmatrix}
		0 & \betaab_{n,0,2} & & \\
		\betaab_{n,1,1} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-2,2} \\
		& & \ddots & 0 \\
		& & & \betaab_{n,n,1} \\
	    \end{pmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots
\end{align}

@@

Recall that, for $a, b$ be general parameters, a function $f(x,y)$ defined on $\Omega$ is approximated by its expansion $f(x,y) = \bigPab(x,y)^\top \mathbf{f}$. Recall also that $\Wab(x,y) := w_H^{(a,b)}(x) \: w_P^{(b)}(\frac{y}{\rho(x)})$, where $\rho(x) := (1-x^2)^\half$, $w_H^{(a,b)}(x) := x^a \: (1-x^2)^b$ and $w_P^{(b)}(y) := (1-y^2)^b$ for the half disk case.

@@

We can gain the non-zero entries of the weighted differential operators similarly, by noting that
\begin{align}
	\pddx \Wab(x,y) &= a W^{(a-1, b)}(x,y) - 2bW^{(a+1, b-1)}(x,y) \\
	\pddy \Wab(x,y) &= -2b \: y \: W^{(a, b-1)}(x,y)
\end{align}
and also that
\begin{align}
	\ip< \Wab \Pnkab, W^{(\tilde{a},\tilde{b})} P^{(\tilde{a},\tilde{b})}_{m,j} >_{W^{(-\tilde{a},-\tilde{b})}} = \ip< \Pnkab, P^{(\tilde{a},\tilde{b})}_{m,j} >_\Wab.
\end{align}
